---
title: "Data 603- Project"
author: "Shora Dehkordi, Ryan leeson, Guarav Kumar, Maryam Sarafraz"
date: "03/12/2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---




```{r Libraries, message=FALSE, warning=FALSE, include=FALSE}
library (ggplot2)
library (mosaic)
library (pROC)
library (ROCR)
library (aod)
library (lmtest)
library (readxl)
library (GGally)
library (markdown)
library (knitr)
library (kableExtra)
library (readxl)
library (mctest)
library (lmtest)
library (olsrr)
library (leaps)
library (car)
```




```{r data import,results="hide"}
taxi_data = read.csv ("./taxitrip_sample_df_final.csv")
#head (taxi_data, 4)
#tail (taxi_data, 7)
```



```{r, results="hide"}
#   convert day_of_week to a numerical value
transform (taxi_data, day_of_week = as.numeric (day_of_week))
```

```{r}
#   Filter for weekend
#   Sunday = 1
#   Saturday = 7
taxi_data$weekend = 1
taxi_data$weekend[ taxi_data$day_of_week > 1 & taxi_data$day_of_week < 6] = 0
```


```{r, results="hide"}
#   convert months to a numerical value
transform (taxi_data, months = as.numeric (months))
```

```{r}
#   Filtering for season
taxi_data$season = "Winter"   #   Winter
taxi_data$season[taxi_data$months > 2 & taxi_data$months < 6] = "Spring"   #   Spring
taxi_data$season[taxi_data$months > 5 & taxi_data$months < 9] = "Summer"   #   Summer
taxi_data$season[taxi_data$months > 8 & taxi_data$months < 12] = "Fall"   #   Fall
```

```{r,results="hide"}
transform (taxi_data, hours = as.numeric (hours))
```

```{r}
taxi_data$time_of_day = "Night"   #   Night
taxi_data$time_of_day[taxi_data$hours >= 6 & taxi_data$hours < 12] = "Morning"   #   Morning
taxi_data$time_of_day[taxi_data$hours >= 12 & taxi_data$hours < 18] = "Afternoon"   #   Afternoon
taxi_data$time_of_day[taxi_data$hours >= 18 & taxi_data$hours < 24] = "Evening"   #   Evening
```

```{r, results="hide"}
head (taxi_data, 4)
```


#### Linear model with log transformation

```{r }
#original model
taxi_fulllm_log = lm ( log (fare) ~ factor(payment_type) + factor(company) + avg_miles + avg_minutes + factor(time_of_day) + factor(season) + factor(weekend) + factor(hour_type), data = taxi_data)
```

#### multi-colinearity

```{r vif}
vif (taxi_fulllm_log)
```
avg_miles and avg_minutes are co-linear

```{r, warning= FALSE, message = FALSE}
ggpairs (taxi_fulllm_log, lower = list ( continuous = "smooth_loess", combo = "facethist", discrete = "facetbar", na = "na"), cardinality_threshold = 25)
```


### Model varaible testing

```{r}
taxi_fulllm_log_nomin = lm ( log (fare) ~ factor(payment_type) + factor(company) + avg_miles + factor(time_of_day) + factor(season) + factor(weekend) + factor(hour_type), data = taxi_data)
```

```{r }
taxi_stepw = ols_step_both_p ( taxi_fulllm_log_nomin, pent = 0.05, prem = 0.1, details = FALSE)
```
Hour_type, company, and avg_miles are suggested for the model.

```{r }
taxi_formodel = ols_step_forward_p ( taxi_fulllm_log_nomin, pent = 0.05, details = FALSE)
```
Hour_type, company and avg_miles are suggested for the model.


```{r}
taxi_backmodel = ols_step_backward_p ( taxi_fulllm_log_nomin, prem = 0.05, details = FALSE)
```
Hour_type, company and avg_miles are suggested for the model.


```{r , echo=FALSE}
ks = ols_step_best_subset (taxi_fulllm_log_nomin, details = FALSE)
par(mfrow=c(2,2)) # split the plotting panel into a 2 x 2 grid
plot(ks$cp,type = "o",pch=12, xlab="Number of Variables",ylab= "Cp")
plot(ks$rsq,type = "o",pch=12, xlab="Number of Variables",ylab= "R^2")
#plot(ks$rss, xlab="Number of Variables",ylab= "RMSE")
plot(ks$aic,type = "o",pch=12, xlab="Number of Variables",ylab= "AIC")
plot(ks$adjr,type = "o",pch=12, xlab="Number of Variables",ylab= "Adjusted R^2")
```

```{r }
ks_stat2 = data.frame ( c(1, 2, 3, 4, 5, 6, 7), ks$cp, ks$aic, ks$adjr, ks$rsq)
names (ks_stat2) = c( "Predictors", "CP", "AIC", "Adjusted R^2", "R^2")
ks_stat2
```
Cp (0.08278259) suggests using the six variable model
AIC (-820.8756) suggersts using the four variable model
Adj.rsq (0.8709532) suggests using the six variable model

```{r }
best.subset = regsubsets ( log (fare) ~ factor(payment_type) + factor(company) + avg_miles + factor(time_of_day) + factor(season) + factor(weekend) + factor(hour_type), data = taxi_data, nv = 10)
summary ( best.subset)
reg.summary = summary ( best.subset)
```

Four variables: company, avg_miles, time_of_day, hour_type
Six variables: company, avg_miles, time_of_day, hour_type, payment_type, season


```{r }
summary (taxi_fulllm_log_nomin)
```
company, time_of_day, hour_type, avg_miles are significant


#### Models

```{r }
#   four variable model
taxi_lm_red_4 = lm ( log (fare) ~ factor(company) + avg_miles + factor(time_of_day) + factor(hour_type), data = taxi_data)

#   three variable model
taxi_lm_red_3 = lm ( log (fare) ~ factor(company) + avg_miles + factor(hour_type), data = taxi_data)

#   six variable model
taxi_lm_red_6 = lm ( log (fare) ~ factor(company) + avg_miles + factor(time_of_day) + factor(hour_type) + factor(payment_type) + factor(season), data = taxi_data)
```

```{r }
taxi_fulllm_log = lm ( log (fare) ~ factor(payment_type) + factor(company) + avg_miles + avg_minutes + factor(time_of_day) + factor(season) + factor(weekend) + factor(hour_type), data = taxi_data)
```

#### Partial F-test
```{r}
#   full and 6 variables 
anova (taxi_fulllm_log_nomin, taxi_lm_red_6)

#   full and 4 variables
anova (taxi_fulllm_log_nomin, taxi_lm_red_4)

#   full and 3 variables
anova (taxi_fulllm_log_nomin, taxi_lm_red_3)
```

```{r}
summary (taxi_lm_red_4)
```

```{r}
anova (taxi_fulllm_log_nomin, taxi_lm_red_3)
```
The F-tests suggest that the three variable model is not significantly different from the full model. So, four variables can be removed from the model.


```{r}
summary (taxi_lm_red_3)
```

```{r}
taxi_lm_red_2 = lm ( log (fare) ~ avg_miles + factor(hour_type), data = taxi_data)
anova (taxi_fulllm_log_nomin, taxi_lm_red_2)
```
Company must stay in the model because p-value < 0.05


So, the three variable model is the best reduced model.


```{r }
print ("Adj. R2")
summary (taxi_lm_red_3)$adj.r.sq
print ("RMSE")
sigma (taxi_lm_red_3)
```



#### Interactions

```{r}
taxi_lm_red_3_int = lm ( log (fare) ~ (factor(company) + avg_miles + factor(hour_type)) ^2, data = taxi_data)
summary (taxi_lm_red_3_int)
```

```{r}
anova (taxi_lm_red_3_int, taxi_lm_red_3)
```
At least one of the interactions are significant.

The individual t-tests suggest only the interaction between avg_miles and company is significant.

```{r}
taxi_lm_red_3_int_red = lm ( log (fare) ~ factor(company) + avg_miles + factor(hour_type) + avg_miles*factor(company), data = taxi_data)
summary (taxi_lm_red_3_int_red)
```

```{r}
anova (taxi_lm_red_3_int_red, taxi_lm_red_3_int)
```
The partial F-test suggests the interactions for company and hour_type, and hour_type and avg_miles are insignificant.

```{r}
anova (taxi_lm_red_3_int_red, taxi_lm_red_3)
```
The partial F-test suggests the interaction between avg_miles and company is significant.

```{r}
summary (taxi_lm_red_3_int_red)
```

```{r }
print ("Adj. R2")
summary (taxi_lm_red_3_int_red)$adj.r.sq
print ("RMSE")
sigma (taxi_lm_red_3_int_red)
```


#### Higher Orders

```{r}
taxi_lm_red_3_int_red_12o = lm ( log (fare) ~ factor(company) + poly (avg_miles, 12, raw = TRUE) + factor(hour_type) + avg_miles*factor(company), data = taxi_data)
summary (taxi_lm_red_3_int_red_12o)
```


```{r}
taxi_lm_red_3_int_red_8o = lm ( log (fare) ~ factor(company) + poly (avg_miles, 8, raw = TRUE) + factor(hour_type) + avg_miles*factor(company), data = taxi_data)
summary (taxi_lm_red_3_int_red_8o)
```


```{r}
taxi_lm_red_3_int_red_7o = lm ( log (fare) ~ factor(company) + poly (avg_miles, 7, raw = TRUE) + factor(hour_type) + avg_miles*factor(company), data = taxi_data)
summary (taxi_lm_red_3_int_red_7o)
```


```{r}
taxi_lm_red_3_int_red_2o = lm ( log (fare) ~ factor(company) + poly (avg_miles, 2, raw = TRUE) + factor(hour_type) + avg_miles*factor(company), data = taxi_data)
summary (taxi_lm_red_3_int_red_2o)
```

```{r}
taxi_lm_red_3_int_red_4o = lm ( log (fare) ~ factor(company) + poly (avg_miles, 4, raw = TRUE) + factor(hour_type) + avg_miles*factor(company), data = taxi_data)
summary (taxi_lm_red_3_int_red_4o)
```

```{r}
anova (taxi_lm_red_3_int_red_8o, taxi_lm_red_3_int_red_12o)
```

```{r }
print ("Adj. R2")
summary (taxi_lm_red_3_int_red)$adj.r.sq
print ("RMSE")
sigma (taxi_lm_red_3_int_red)
```


#### Test of assumptions

```{r }

#   2nd order model
ggplot (taxi_lm_red_3_int_red_2o, aes ( x = .fitted, y = .resid)) +
  geom_point () + geom_smooth () +
  geom_hline (yintercept = 0)

#   12th order model
ggplot (taxi_lm_red_3_int_red_12o, aes ( x = .fitted, y = .resid)) +
  geom_point () + geom_smooth () +
  geom_hline (yintercept = 0)
```


```{r }
ggplot (taxi_lm_red_3_int_red_12o, aes ( x = .fitted, y = sqrt ( abs (.stdresid)))) +
  geom_point () + geom_smooth () +
  geom_hline (yintercept = 0) + 
  ggtitle ("Scale-Location plot: Standardised Residual vs Fitted values")


#   BP test
bptest (taxi_lm_red_3_int_red_12o)
#   H0 : heteroscedasticity is not present

```


```{r }
par ( mfrow = c(1,2))
ggplot ( data = taxi_data, aes ( residuals (taxi_lm_red_3_int_red_12o))) + 
  geom_histogram (breaks = seq (-1, 1, by = 0.1), col = "red", fill = "blue") + 
  labs ( title = "Histogram for residuals") +
  labs ( x = "residuals", y = "Count")

ggplot (taxi_data, aes ( sample = taxi_lm_red_3_int_red_12o$residuals)) + 
  stat_qq () +
  stat_qq_line ()

```


```{r }
shapiro.test ( residuals (taxi_lm_red_3_int_red_12o))
#   H0 : model is normal
```


#### Test for outliers 

```{r }
# order 12 cook's distance for 12th order model
plot (taxi_lm_red_3_int_red_12o, which = 5)
```


```{r }
plot (taxi_lm_red_3_int_red_2o, which = 5)
```


```{r }
taxi_data[cooks.distance (taxi_lm_red_3_int_red_12o) > 0.5,]
```

```{r}
plot (taxi_lm_red_3_int_red_12o, pch = 10, col = "red", which = c(4))
```


```{r }
lev = hatvalues (taxi_lm_red_3_int_red_12o)
p = length ( coef (taxi_lm_red_3_int_red_12o))
n = nrow (taxi_data)
outlier = lev[lev > (2*p/n)]
print (outlier)
```

```{r }
lev = hatvalues (taxi_lm_red_3_int_red_12o)
p = length ( coef (taxi_lm_red_3_int_red_12o))
n = nrow (taxi_data)
outlier = lev[lev > (3*p/n)]
print (outlier)
```

(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,263,288,300,401,408,419,436,437,528,609,613,615,1081,1293,1525,1555,1562,1571,1578,1586,1599,1862,1867,1954,2117,2122,2142,2143,2149,2268,2314,2577,2665,2707,2711,2722,2725,2733,2751,2752,2753,2754,2755,2756,2757,2758,2759,2760,2761,2762,2763,2764,2765,2766,2767,2768,27692770,2771,2772,2773,2774,2775,2776,2777,2778,2779,2780,2781,2782,2783,2784,2785,2786,2787,2788,2789,2790,2791,2792,2793,2794,2795,2796,2797,2798,2799,2800,2901,3034,3182,3269,3301,3305,3306,3320,3361,3445,3578,3727,4012,4573,4652,4673)

```{r }
plot (rownames (taxi_data), lev, main = "Leverage in taxi dataset", xlab = "observation", ylab = "Leverage Value") 
abline (h = 2*p/n, lty = 1)
abline (h = 3*p/n, lty = 1)
```


### New dataset

```{r }
taxi_data2 = taxi_data[-c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,263,288,300,401,408,419,436,437,528,609,613,615,1081,1293,1525,1555,1562,1571,1578,1586,1599,1862,1867,1954,2117,2122,2142,2143,2149,2268,2314,2577,2665,2707,2711,2722,2725,2733,2751,2752,2753,2754,2755,2756,2757,2758,2759,2760,2761,2762,2763,2764,2765,2766,2767,2768,27692770,2771,2772,2773,2774,2775,2776,2777,2778,2779,2780,2781,2782,2783,2784,2785,2786,2787,2788,2789,2790,2791,2792,2793,2794,2795,2796,2797,2798,2799,2800,2901,3034,3182,3269,3301,3305,3306,3320,3361,3445,3578,3727,4012,4573,4652,4673), ]
```

```{r}
nrow (taxi_data2)
```

```{r  }
taxi2_fulllm_log = lm ( log (fare) ~ factor(payment_type) + factor(company) + avg_miles + avg_minutes + factor(time_of_day) + factor(season) + factor(weekend) + factor(hour_type), data = taxi_data2)
```

```{r }
vif (taxi2_fulllm_log)
```
avg_minutes should be removed

### Model varaible testing

```{r }
taxi2_fulllm_log_nomin = lm ( log (fare) ~ factor(payment_type) + factor(company) + avg_miles + factor(time_of_day) + factor(season) + factor(weekend) + factor(hour_type), data = taxi_data2)
```

```{r }
taxi_stepw = ols_step_both_p ( taxi2_fulllm_log_nomin, pent = 0.05, prem = 0.1, details = FALSE)
```
Stepwise regression suggests a model including avg_miles, hour_type, company, and time_of_day.

```{r }
taxi_formodel = ols_step_forward_p ( taxi2_fulllm_log_nomin, pent = 0.05, details = FALSE)
```
Forward regression suggests a model including avg_miles, hour_type, company, and time_of_day. 

```{r }
taxi_backmodel = ols_step_backward_p ( taxi2_fulllm_log_nomin, prem = 0.05, details = FALSE)
```
Backward regression suggests a model including avg_miles, hour_type, company, and time_of_day.

```{r echo=FALSE, cache=TRUE}
ks = ols_step_best_subset (taxi2_fulllm_log_nomin, details = FALSE)
par(mfrow=c(2,2)) # split the plotting panel into a 2 x 2 grid
plot(ks$cp,type = "o",pch=12, xlab="Number of Variables",ylab= "Cp")
plot(ks$rsq,type = "o",pch=12, xlab="Number of Variables",ylab= "R^2")
#plot(ks$rss, xlab="Number of Variables",ylab= "RMSE")
plot(ks$aic,type = "o",pch=12, xlab="Number of Variables",ylab= "AIC")
plot(ks$adjr,type = "o",pch=12, xlab="Number of Variables",ylab= "Adjusted R^2")
```

```{r }
ks_stat2 = data.frame ( c(1, 2, 3, 4, 5, 6, 7), ks$cp, ks$aic, ks$adjr, ks$rsq)
names (ks_stat2) = c( "Predictors", "CP", "AIC", "Adjusted R^2", "R^2")
ks_stat2
```
Cp suggests using the six variable model
AIC suggersts using the four variable model
Adj.rsq suggests using the six variable model

```{r }
best.subset = regsubsets ( log (fare) ~ factor(payment_type) + factor(company) + avg_miles + factor(time_of_day) + factor(season) + factor(weekend) + factor(hour_type), data = taxi_data2, nv = 10)
summary ( best.subset)
reg.summary = summary ( best.subset)
```

```{r }
summary (taxi2_fulllm_log_nomin)
```
Season, weekend, payment_type are insignificant.

#### Models
```{r}
taxi2_lm_red_4 = lm ( log (fare) ~ factor(company) + avg_miles + factor(time_of_day) + factor(hour_type), data = taxi_data2)
taxi2_lm_red_3 = lm ( log (fare) ~ factor(company) + avg_miles + factor(hour_type), data = taxi_data2)

taxi2_lm_red_6 = lm ( log (fare) ~ factor(company) + avg_miles + factor(time_of_day) + factor(hour_type) + factor(payment_type) + factor(season), data = taxi_data2)
```

```{r}
taxi2_fulllm_log = lm ( log (fare) ~ factor(payment_type) + factor(company) + avg_miles + avg_minutes + factor(time_of_day) + factor(season) + factor(weekend) + factor(hour_type), data = taxi_data2)
```

```{r}
nrow (taxi_data2)
```

#### Partial F-test
```{r}
#   full and 6 variables 
anova (taxi2_fulllm_log_nomin, taxi2_lm_red_6)

#   full and 4 variables
anova (taxi2_fulllm_log_nomin, taxi2_lm_red_4)

#   full and 3 variables
anova (taxi2_fulllm_log_nomin, taxi2_lm_red_3)
```

```{r}
summary (taxi2_lm_red_4)
```

```{r}
anova (taxi2_lm_red_4, taxi2_lm_red_3)
```
time_of_day is significant according to the above results, so, the four variable model is selected.


```{r }
print ("Adj. R2")
summary (taxi2_lm_red_4)$adj.r.sq
print ("RMSE")
sigma (taxi2_lm_red_4)
```


#### Interactions

```{r}
taxi2_lm_red_4_int = lm ( log (fare) ~ (factor(company) + avg_miles + factor(hour_type) + factor(time_of_day)) ^2, data = taxi_data2)
summary (taxi2_lm_red_4_int)
```
From the individual t-tests, company*hour_type appears to be insignificant.

```{r}
anova (taxi2_lm_red_4_int, taxi2_lm_red_4)
```
The f-test suggests that the interactions are significant

```{r}
#   Interaction model without compnay*time_of_day
taxi2_lm_red_4_int_red = lm ( log (fare) ~ factor(company) + avg_miles + factor(hour_type) + factor(time_of_day) + avg_miles*factor(company) + factor(hour_type)*factor(time_of_day) + avg_miles*factor(time_of_day), data = taxi_data2)


#   Partial F-test
anova (taxi2_lm_red_4_int_red, taxi2_lm_red_4_int)
```
The partial F-test indicates company*time_of_day is an insignificant interaction (F= 0.7544, df= 9, 4824, p-value = 0.659)


```{r}
#   Interaction model without hour_type*time_of_day
taxi2_lm_red_4_int_red_2 = lm ( log (fare) ~ factor(company) + avg_miles + factor(hour_type) + factor(time_of_day) + avg_miles*factor(company) + avg_miles*factor(time_of_day), data = taxi_data2)

#   Partial F-Test
anova (taxi2_lm_red_4_int_red_2, taxi2_lm_red_4_int_red)
```
The partial F-test indicates hour_type*time_of_day must be kept (F= 10.878, df= 2, 4833, p-value < 0.05)


```{r}
#   Interaction model without avg_miles*time_of_day
taxi2_lm_red_4_int_red_3 = lm ( log (fare) ~ factor(company) + avg_miles + factor(hour_type) + factor(time_of_day) + avg_miles*factor(company) + factor(hour_type)*factor(time_of_day), data = taxi_data2)


#   Partial F-test
anova (taxi2_lm_red_4_int_red_3, taxi2_lm_red_4_int_red)
```
The partial F-test indicates avg_miles*time_of_day must be kept (F= 4.3922, df= 3, 4833, p-value= 0.004303 < 0.05)

```{r}
#   Interaction model without avg_mlies*company 
taxi2_lm_red_4_int_red_4 = lm ( log (fare) ~ factor(company) + avg_miles + factor(hour_type) + factor(time_of_day) + factor(hour_type)*factor(time_of_day) + avg_miles*factor(time_of_day), data = taxi_data2)

#   Partial F-test 
anova (taxi2_lm_red_4_int_red_4, taxi2_lm_red_4_int_red)
```
The partial F-test indicates avg_miles*company must be kept (F= 6.405, df= 2, 4833, p-value = 0.001667 < 0.05)


```{r }
taxi2_lm_red_4_int_red_2o = lm ( log (fare) ~ factor(company) + poly (avg_miles, 2, raw = TRUE) + factor(hour_type) + factor(time_of_day) + avg_miles*factor(company) + factor(hour_type)*factor(time_of_day) + avg_miles*factor(time_of_day), data = taxi_data2)
summary (taxi2_lm_red_4_int_red_2o)
```


```{r 9th order model}
taxi2_lm_red_4_int_red_9o = lm ( log (fare) ~ factor(company) + poly (avg_miles, 9, raw = TRUE) + factor(hour_type) + factor(time_of_day) + avg_miles*factor(company) + factor(hour_type)*factor(time_of_day) + avg_miles*factor(time_of_day), data = taxi_data2)
summary (taxi2_lm_red_4_int_red_9o)
```

```{r 15 th order model}
taxi2_lm_red_4_int_red_15o = lm ( log (fare) ~ factor(company) + poly (avg_miles, 15, raw = TRUE) + factor(hour_type) + factor(time_of_day) + avg_miles*factor(company) + factor(hour_type)*factor(time_of_day) + avg_miles*factor(time_of_day), data = taxi_data2)
summary (taxi2_lm_red_4_int_red_15o)
```

```{r }
print ("Adj. R2")
summary (taxi2_lm_red_4_int_red_9o)$adj.r.sq
print ("RMSE")
sigma (taxi2_lm_red_4_int_red_9o)
```


#### Test of assumptions

```{r }
ggplot (taxi2_lm_red_4_int_red_9o, aes ( x = .fitted, y = .resid)) +
  geom_point () + geom_smooth () +
  geom_hline (yintercept = 0) + 
  ggtitle ("9th order")


ggplot (taxi2_lm_red_4_int_red_2o, aes ( x = .fitted, y = .resid)) +
  geom_point () + geom_smooth () +
  geom_hline (yintercept = 0) + 
  ggtitle ("2nd order")
```

```{r }
ggplot (taxi2_lm_red_4_int_red_9o, aes ( x = .fitted, y = sqrt ( abs (.stdresid)))) +
  geom_point () + geom_smooth () +
  geom_hline (yintercept = 0) + 
  ggtitle ("Scale-Location plot: Standardised Residual vs Fitted values, 9th order")

ggplot (taxi2_lm_red_4_int_red_2o, aes ( x = .fitted, y = sqrt ( abs (.stdresid)))) +
  geom_point () + geom_smooth () +
  geom_hline (yintercept = 0) + 
  ggtitle ("Scale-Location plot: Standardised Residual vs Fitted values, 2th order")


#   Bp test, 9th order model
bptest (taxi2_lm_red_4_int_red_9o)
#   H0 : heteroscedasticity is not present

#   Bp test, 2nd order model
bptest (taxi2_lm_red_4_int_red_2o)

```


```{r }
ggplot ( data = taxi_data2, aes ( residuals (taxi2_lm_red_4_int_red_9o))) + 
  geom_histogram (breaks = seq (-1, 1, by = 0.1), col = "red", fill = "blue") + 
  labs ( title = "Histogram for residuals") +
  labs ( x = "residuals", y = "Count")


ggplot (taxi_data2, aes ( sample = taxi2_lm_red_4_int_red_9o$residuals)) + 
  stat_qq () +
  stat_qq_line ()


shapiro.test ( residuals (taxi2_lm_red_4_int_red_9o))
#   H0 : model is normal

```


```{r }
ggplot ( data = taxi_data2, aes ( residuals (taxi2_lm_red_4_int_red_2o))) + 
  geom_histogram (breaks = seq (-1, 1, by = 0.1), col = "red", fill = "blue") + 
  labs ( title = "Histogram for residuals") +
  labs ( x = "residuals", y = "Count")


ggplot (taxi_data2, aes ( sample = taxi2_lm_red_4_int_red_2o$residuals)) + 
  stat_qq () +
  stat_qq_line ()


shapiro.test ( residuals (taxi2_lm_red_4_int_red_2o))
#   H0 : model is normal

```


#### Outliers

```{r}
plot (taxi2_lm_red_4_int_red_2o, which = 5)

plot (taxi2_lm_red_4_int_red_9o, which = 5)
```



```{r }
taxi_data[cooks.distance (taxi2_lm_red_4_int_red_9o) > 0.5,]

taxi_data[cooks.distance (taxi2_lm_red_4_int_red_2o) > 0.5,]
```


```{r}
plot (taxi2_lm_red_4_int_red_9o, pch = 10, col = "red", which = c(4))

plot (taxi2_lm_red_4_int_red_2o, pch = 10, col = "red", which = c(4))
```

```{r }
lev = hatvalues (taxi2_lm_red_4_int_red_9o)
p = length ( coef (taxi2_lm_red_4_int_red_9o))
n = nrow (taxi_data2)
outlier = lev[lev > (3*p/n)]
print (outlier)
```

```{r }
plot (rownames (taxi_data2), lev, main = "Leverage in taxi dataset", xlab = "observation", ylab = "Leverage Value") 
abline (h = 2*p/n, lty = 1)
abline (h = 3*p/n, lty = 1)
```




